Model to study co-evolution of trade an culture

### 2015-03-25 article.tex
integrate corrections of bib to zotero, add new articles from zotero to bib, update economy state of the art

\usepackage{xcolor}

@@

\title{Model to study co-evolution of trade an culture}

\begin{figure*}[htb]
{
\centering
Jean-Marc Montanier\\
Simon Carrignon\\ 
Xavier Rubio\\
\vspace{12pt}
Barcelona Supercomputing Center\\
Carrer de Jordi Girona, 29, \\
08034 Barcelona, Spain\\
}
\end{figure*}

\maketitle

@@

Multiple approaches have been proposed to model the economical interactions in the ancient mediteranean culture. Among these we will focus on the formal models. These are close to our works, can be tested and compared.

%Hopkins 1980: structure of tax system. two hypothesis. 1)taxation increased volume of trade in the empire? 2) tax-exporting province (recently integrated province) had to export to pay tax. roman supported by many small taxes. Not sure to cite it, goes to much more
%Termin 2001: based on multiple consideration classify roman economy as market based with loose tie from things far appart. composed of local markets connected together
%Terpstra 2011: micro economic model to explain possibility of trade in roman empire under uncertainty. based on reputation in the network and removal of bad from the network (personalized network based on geographical info). focus on a specific case (one city and its relations with others). threat of expulsion or loss of reputation as reinforcement <- not only economic.

Finally we would like to add this additional work even if it has not been specifically designed for historical studies~\cite{gintis_emergence_2006}. This model proposes a specific analysis on the emergence of price structure when only pair exchanges are considered, thus observing the emergence of a market economy. This constitutes a relatively realistic model with simple assumptions and will be explained in further details in section XXX.

@@

\xrc{what is exactly the definition of culture used in the paper? I would suggest this one: \emph{Culture is socially learned information capable of affecting  individual phenotypes. People acquire culture from other  individuals, via teaching or imitation} \cite{richerson1996}}

The second aspect we wish to model is the evolution of the culture in the Roman empire. The evolutionary framework (generally used in Biology) has been proposed to study the changes of culture and found to be suitable,  particularly in the case of historical and archaeological studies~\cite{lycett_cultural_2015,henrich_evolution_2003}\xrc{henrich's paper is not on archaeology or history, better reference \cite{shennan_2008}}.

@@

The variation mechanisms are at the heart of the creation of the creation of new cultural elements~\cite{obrien_variation_1990}. These may happen due to some errors during the copy process~\cite{schillinger_copying_2014}. In more recent culture we can witness the apparition of mutations with intentionality~\cite{ziman_technological_2003}\xrc{see also recombination \cite{sole_2013}}.

\xrc{Instead of differential replication I would use \emph{biased selection}}
The last element is the one that will interest us the most in this study: the differential replication between different cultural traits. If a cultural trait is found more often than the others in the population it means that something modified the rate at which it is adopted by each agent \xrc{beyond drift effects produced by random copy of neutral traits \cite{bentley_2004}}.
%The reproduction rate of ideas in a population can be characterise by different rates: directional, disruptive or stabilisation~\cite{rendell_cognitive_2011}. 
Multiple processes can be at play behind the reproduction \xrc{selection? transmission? reproduction is too biological} of cultural traits. Among them the three following will interest us due to their simplicity to test in an agent-based modelisation. First is the random copy of one trait from an agent to another agent. 
\todo{get a ref on that} \xrc{you can use \cite{bentley_2004}}
This type of mechanism reproduces well the changes in musical preferences \xrc{music here is odd; if an example is used it should be related to the topic of the paper}. Second are the the copy based on the frequency at which a cultural trait is observed. If a common trait is reproduced more often then we speak about ``conformity.'' If on the contrary the less common trait is preferred for reproduction we speak of ``rarity'' \xrc{nope, we speak of anti-conformism}. Finally a trait can be preferred depending on whom is carrying it. Among these preferences the prestige of the individual carrying it can constitute an explanation for the adoption foreign cultural traits in the Roman empire \xrc{this is known as model biased transmission}.

@@

This model is composed of multiple agents producing goods of a fixed number of types. The agents are endowed with storage abilities for each good that can be produced. On top of	producing the agents have the ability to exchange and consume the goods. The specific rules for production, consumption and trading will vary depending on the hypothesis tested. Nevertheless the structure of the environment and the proceeding of a simulation will be the same for all hypothesis tested in this model.

@@

More formally the model is composed as follow. $r$ is the number of ressources present in the environment. The whole population is defined by a vector of $m$ agents: 
		$$ population = (a_i, \cdots ,a_m) $$

Each agent makes its own estimate on the price to attribute to a certain good. This price estimate is used for both selling and buying goods :
$$ price = (p^i_1,\cdots,p^i_n) $$

@@

The elments listed above constitute the basic structure of the model. Different behaviours can be used for the agents and world depending on the assumption to test. This scheme allow us to have a model ahieving the two main goals listed in section~\ref{}. First, this model has a simple behaviour and few parameters to analyse. Second, this model is flexible enough to allow the study of multiple hypothesis while conserving the same units that can therefore be compared between simulations.

\subsection{Validation}
\xrc{Validation is impossible for a theoretical model. Either you use verification if the goal is to check that implementation is correct, or design experiment for defining the experimens that will be used to explore the model.}

>>>>

\usepackage{xcolor}
    \usepackage{cite}

@@

\title{Model to study co-evolution of trade an culture}
\maketitle

\begin{figure*}[htb]
{
\centering
Jean-Marc Montanier\\
Simon Carrignon\\ 
Xavier Rubio-Campillo\\
\vspace{12pt}
Barcelona Supercomputing Center\\
Carrer de Jordi Girona, 29, \\
08034 Barcelona, Spain\\
}
\end{figure*}

@@

%Hopkins 1980: structure of tax system. two hypothesis. 1)taxation increased volume of trade in the empire? 2) tax-exporting province (recently integrated province) had to export to pay tax. roman supported by many small taxes. Not sure to cite it

Within the archaeological field a large number of works have been proposed to quantify the properties of the economical interactions in the ancient Mediterranean societies~\cite{hopkins_taxes_1980,temin_market_2001,terpstra_trade_2011,minc_monitoring_2006,temin_economy_2006,wilson_approaches_2009,scheidel_model_2007,kessler_organization_2007}. In the works of~\cite{temin_market_2001,temin_economy_2006,wilson_approaches_2009}, based on the analysis of archaeological evidence, the Roman economy has been classified as a type of market economy where a variety of small tightly connected markets had loose connections between them. This view of a market system in ancient societies has been further expended in~\cite{minc_monitoring_2006} by a model applicable to various cases and studied in a test case of the aztec economical structure. The reaming works attempted at defining more exactly which type market economy it was, i.e. was it close to ours or remote ?

The work of~\cite{scheidel_model_2007} proposes to analyse the factors affecting the growth of income and population in the Roman Italy. The analysis of historical evidence and historical documents lead the author to propose a model highlighting the key factors in the economical growth. More in link with the work conducted in this article~\cite{terpstra_trade_2011,kessler_organization_2007} have analysed the interactions between economy and culture transmission in roman empire. An analysis based on historical documents in~\cite{kessler_organization_2007} highlights the importance of economical and social institutions in order to perform an efficient trade of trains in the early Roman Empire. More precisely, the work conducted in~\cite{terpstra_trade_2011} proposes a micro-economic model analysing the possibility of trade under uncertainty in the roman empire. This work highlight the importance of reputation and the threat of expulsion from a trade network as mechanisms the overall efficiency of the trade network.

A limitation of the works cited above is the impossibility to test the conclusions reached by the authors. The models proposed reflects only the view of the authors on the data they have analyse. A new type of model would be necessary to go beyond the interpretation, and allow the verification of the conclusions reached. These models should be able to generate data which is then qualitatively compared to the real historical data. By this mean, if a factor is really important in the economical network, its removal from the model will lead to data qualitatively different from the historical data. The Multi-Agent Simulations (MAS) offer a way to produce this kind of model~\cite{lake_trends_2014}.

Based on this tool a relatively complex and exact modelisation of the prehispanic Pueblo societies has been proposed~\cite{kohler_modelling_2012}. This work offer a relatively exact MAS integrating notable the effects food consumption, production and landscape. The questions studied are  numerous and some works are performed on the economical system and its ties with social transmissions~\cite{kobti_emergence_2006,cockburn_simulating_2013}. However the work done in this project consider a pre-market society in a specific village which makes it difficult to reach more general conclusions, notably applicable to the Roman Empire.

%kobti 2012 la meme que kobti 2006 ?
%	pour 2006: etude de l'apparition d'inegalite

@@

\xrc{what is exactly the definition of culture used in the paper? I would suggest this one: \emph{Culture is socially learned information capable of affecting  individual phenotypes. People acquire culture from other individuals, via teaching or imitation} \cite{richerson_principles_1996}}

The second aspect we wish to model is the evolution of the culture in the Roman empire. The evolutionary framework (generally used in Biology) has been proposed to study the changes of culture and found to be suitable,  particularly in the case of historical and archaeological studies~\cite{lycett_cultural_2015,henrich_evolution_2003}\xrc{henrich's paper is not on archaeology or history, better reference \cite{shennan_evolution_2008}}.

@@

The variation mechanisms are at the heart of the creation of the creation of new cultural elements~\cite{obrien_variation_1990}. These may happen due to some errors during the copy process~\cite{schillinger_copying_2014}. In more recent culture we can witness the apparition of mutations with intentionality~\cite{ziman_technological_2003}\xrc{see also recombination \cite{sole_evolutionary_2013}}.

\xrc{Instead of differential replication I would use \emph{biased selection}}
The last element is the one that will interest us the most in this study: the differential replication between different cultural traits. If a cultural trait is found more often than the others in the population it means that something modified the rate at which it is adopted by each agent \xrc{beyond drift effects produced by random copy of neutral traits \cite{bentley_random_2004}}.
%The reproduction rate of ideas in a population can be characterise by different rates: directional, disruptive or stabilisation~\cite{rendell_cognitive_2011}. 
Multiple processes can be at play behind the reproduction \xrc{selection? transmission? reproduction is too biological} of cultural traits. Among them the three following will interest us due to their simplicity to test in an agent-based modelisation. First is the random copy of one trait from an agent to another agent. 
\todo{get a ref on that} \xrc{you can use \cite{bentley_random_2004}}
This type of mechanism reproduces well the changes in musical preferences \xrc{music here is odd; if an example is used it should be related to the topic of the paper}. Second are the the copy based on the frequency at which a cultural trait is observed. If a common trait is reproduced more often then we speak about ``conformity.'' If on the contrary the less common trait is preferred for reproduction we speak of ``rarity'' \xrc{nope, we speak of anti-conformism}. Finally a trait can be preferred depending on whom is carrying it. Among these preferences the prestige of the individual carrying it can constitute an explanation for the adoption foreign cultural traits in the Roman empire \xrc{this is known as model biased transmission}.

@@

The only mandatory elements are a set of agents each possessing a fixed number of goods and associate a value to each of these goods. The value can be for example either an interest in the good or the price associated to this good. The agents are also endowed with limited storage abilities for each good that can be produced.

More formally the model is composed as follow. $r$ is the number of resources present in the environment. The whole population is defined by a vector of $m$ agents: 
$$ population = (a_i, \cdots ,a_m) $$

Each agent makes its own estimate on the price to attribute to a certain good. This price estimate is used for both selling and buying goods :
$$ values = (v^i_1,\cdots,p^i_n) $$

On top of producing the agents have the ability to exchange and consume the goods. 
On the one side, the structure of the environment and the proceeding of a simulation will be the same for all hypothesis tested in this model. The specific rules for production, consumption and trading vary depending on the hypothesis tested. Nevertheless

@@

@@

The elements listed above constitute the basic structure of the model. Different behaviours can be used for the agents and world depending on the assumption to test. This scheme allow us to have a model achieving the two main goals listed in section~\ref{}. First, this model has a simple behaviour and few parameters to analyse. Second, this model is flexible enough to allow the study of multiple hypothesis while conserving the same units that can therefore be compared between simulations.

\subsection{Validation}
\xrc{Validation is impossible for a theoretical model. Either you use verification if the goal is to check that implementation is correct, or design experiment for defining the experiments that will be used to explore the model.}
